\documentclass[10pt]{sig-alternate-05-2015}
%\documentclass{sig-alternate-10pt}
%\documentclass{sigcomm-alternate}

\usepackage[l2tabu,orthodox]{nag}  % Check for use of obsolete LaTeX packages/constructs
\usepackage[utf8x]{inputenc}       % This file is formatted in UTF-8
\usepackage{upquote}               % Fix quotes in verbatim environments

\usepackage{graphicx}              %
\usepackage{amsmath}               % http://ctan.org/pkg/amsmath
%% this breaks tikz... badly...
%\usepackage[all]{onlyamsmath}      % http://ctan.org/tex-archive/macros/latex/contrib/onlyamsmath
\usepackage{url}                   % http://ctan.org/tex-archive/macros/latex/contrib/url
\usepackage[caption=false]{subfig} % http://ctan.org/tex-archive/macros/latex/contrib/subfig
                                   % The subfig package is deprecated in favour of subcaption,
                                   % but as of April 2015 subcaption doesn't work with ACM or
                                   % IEEE style files (this is also the reason for [caption=false])

\usepackage{verbatim, xcolor}
\usepackage{enumitem}
\usepackage{acronyms}
\usepackage{natbib}
\def\bibfont{\small}
\def\bibsep{0pt}
\usepackage{paralist}
\usepackage{whix} %% shiny wcix diagrams
\usepackage{todonotes}
\newcommand{\narrative}[1]{\textcolor{blue}{(narrative: #1)}}

%%MF -- To put all authors on front page, and save space.
\newcommand{\superscript}[1]{\raisebox{1ex}{\ensuremath{{\mathrm{#1}}}}}
%\def\sharedaffiliation{\end{tabular}\newline\begin{tabular}{c}}
\def\we{\superscript{\ast}}
\def\wh{\superscript{\dag}}
\def\wg{\superscript{\ddag}}
\def\ws{\superscript{\natural}}

\begin{document}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}


% DOI
\doi{http://dx.doi.org/10.1145/2940157.2940162}

% ISBN
\isbn{978-1-4503-4423-4/16/08}

% Conference
\conferenceinfo{GAIA,}{August 22-26, 2016, Florianopolis, Brazil}
\CopyrightYear{2016}

\acmPrice{\$15.00}


% --- Author Metadata here ---
% \conferenceinfo{GAIA,}{August 22-26, 2016 Florianopolis, Brazil}
% \CopyrightYear{2016} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
% \crdata{978-1-4503-4423-4}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{RemIX: A Distributed Internet Exchange\\for Remote and Rural Networks
% \titlenote{(Produces the permission block, and
% copyright information). For use with
% SIG-ALTERNATE.CLS. Supported by ACM.}
}
% \subtitle{(Pronounced 'kicks,' for Community Internet Exchange)
% \titlenote{A full version of this paper is available as
% \textit{Author's Guide to Preparing ACM SIG Proceedings Using
% \LaTeX$2_\epsilon$\ and BibTeX} at
% \texttt{www.acm.org/eaddress.htm}}
%}

%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
 \alignauthor
  William Waites\we\wh, James Sweet\we\wh, Roger Baig\wg,\\
  Peter Buneman\we, Marwan Fayed\ws, Gordon Hughes\we\\
 \affaddr{{\we}University of Edinburgh, {\wh}HUBS \textsc{c.i.c},
   {\wg}guifi.net, {\ws}University of Stirling}\\
 {\email{\small Corresponding authors: wwaites@tardis.ed.ac.uk, opb@inf.ed.ac.uk, mmf@cs.stir.ac.uk}}
% \email{wwaites@tardis.ed.ac.uk, \{opb,mfourman\}@inf.ed.ac.uk, james@hubs.net.uk, roger.baig@guifi.net, mmf@cs.stir.ac.uk, r.a.simmons@stir.ac.uk}
%  William Waites \\ %\titlenote{Dr.~Trovato insisted his name be first.}\\
% %        \affaddr{School of }\\
% %        \affaddr{1932 Wallamaloo Lane}\\
% %        \affaddr{Wallamaloo, New Zealand}\\
% %        \email{trovato@corporation.com}
% % 2nd. author
%  \alignauthor
%  James Sweet \\
%  \alignauthor
%  Roger Baig \\
%  \and
%  \alignauthor
%  Peter Buneman  %\titlenote{The secretary disavows
% % any knowledge of this author's actions.}\\
% %        \affaddr{Institute for Clarity in Documentation}\\
% %        \affaddr{P.O. Box 1212}\\
% %        \affaddr{Dublin, Ohio 43017-6221}\\
% %        \email{webmaster@marysville-ohio.com}
% % % 3rd. author
%  \alignauthor
%  Marwan Fayed \\ %\titlenote{This author is the
% % one who did all the really hard work.}\\
% %        \affaddr{The Th{\o}rv{\"a}ld Group}\\
% %        \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
% %        \affaddr{Hekla, Iceland}\\
% %        \email{larst@affiliation.org}
% % use '\and' if you need 'another row' of author names
% % % 4th. author
%  \alignauthor
%  Gordon Hughues \\
% %        \affaddr{Brookhaven Laboratories}\\
% %        \affaddr{Brookhaven National Lab}\\
% %        \affaddr{P.O. Box 5000}\\
% %        \email{lleipuner@researchlabs.org}
}
\additionalauthors{Additional authors: Roger Baig (guifi.net), Michael Fourman (University of
Edinburgh), and Richard Simmons (University of Stirling)
% email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
% (The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).
}

\maketitle

\begin{abstract}

The \ac{IXP}, an Ethernet fabric central to the
structure of the global Internet, is largely absent from
% the development of
community-driven collaborative network infrastructure.
%The reasons for this are two-fold.
\acp{IXP} exist in central,
typically urban, environments where strong network infrastructure
ensures high
levels of connectivity.
%Between rural and remote regions,
Between rural and remote networks, separated by distance and terrain,
no such infrastructure exists. In this paper we present RemIX a
distributed \ac{IXP} architecture designed for the community network
environment. We examine this praxis using an implementation in
Scotland, with suggestions for future development and research.

\begin{comment} %MF-comment
The concept of the \ac{IXP} is, so central to the
structure of the global Internet has so far been mostly absent
from the world of community-driven collaborative network
infrastructure development. The dual reasons for this are that
\acp{IXP} typically exist in a central place in an urban area
and community networks are often in rural and remote places,
with long distances, mountains and the sea to contend with.
In this paper we present RemIX a distributed \acp{IXP}
architecture designed for the community network environment
and examine two instances of this praxis in Scotland and
Catalunya.


\par\vspace{\baselineskip}
\todo[inline]{Below deprecated for now in discussion with RB
for purposes of abstract submission deadline.}
Salient differences between HUBS and Guifi:
\begin{itemize}
    \item HUBS lacks a centralising structure, a foundation, with an articulated social contract.
    \item Guifi itself does not provide Internet access
    \item HUBS is an IP Transit provider
    \item If anything, Guifi is more like a (larger version of) WHIX
    \item There is some conflict of interest in HUBS operating WHIX, which Guifi does not suffer from
\end{itemize}
\hrule

Rural access networks are designed to bridge the `last-mile' broadband
gap in regions that are under-served by traditional broadband
providers. Their construction is bespoke, driven by their
beneficiaries, and determined by physical landscape, population
distribution, and monetary budget. Irrespective of their differences,
they are joined by one substantial challenge: connecting to the rest
of the Internet is prohibitively expensive. HUBS \textsc{c.i.c} was
created in Scotland to respond to this. It is a co-operative of access
network members that generates the economies of scale required to
afford backhaul and Internet transit. While intermediation at the IP
layer between the members and the Internet is required for reasons of
scale, it is neither necessary nor desirable amongst the member
networks themselves. In urban areas, networks could interconnect with
each other at an Internet Exchange Point (IXP). In Scotland where the
networks are scattered across 80,000km$^2$ of mountain, field, and sea
it is not so easy. We bridge this gap with a design for a distributed
Internet exchange for access networks in remote places. Doing so
allows for bilateral arrangements for mutual support and assistance
between these networks, and increases the resilience of access network
connectivity. We present the relevant components, and describe our
implementation, so that our efforts may be reproduced.
\end{comment} % MF-comment

\end{abstract}


%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste CCSXML code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
  <concept>
    <concept_id>10003033.10003034</concept_id>
    <concept_desc>Networks~Network architectures</concept_desc>
    <concept_significance>500</concept_significance>
  </concept>
  <concept>
    <concept_id>10003033.10003106.10003119</concept_id>
    <concept_desc>Networks~Wireless access networks</concept_desc>
    <concept_significance>500</concept_significance>
  </concept>
  % <concept>
  %   <concept_id>10003456.10003462.10003561.10003560</concept_id>
  %   <concept_desc>Social and professional topics~Broadband access</concept_desc>
  %   <concept_significance>500</concept_significance>
  % </concept>
  % <concept>
  %   <concept_id>10003456.10003462.10003561.10003566</concept_id>
  %   <concept_desc>Social and professional topics~Universal access
  %   </concept_desc>
  %   <concept_significance>500</concept_significance>
  % </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Networks~Network architectures}
\ccsdesc[500]{Networks~Wireless access networks}
%\ccsdesc[500]{Social and professional topics~Broadband access}
% \ccsdesc[500]{Social and professional topics~Universal access}

\printccsdesc

\keywords{Internet Exchanges (IXP); Community Broadband}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction} \label{sec:intro} \input{intro}

\section{\aclp*{IXP}} \label{sec:context} \input{context}

\section{RemIX Architecture} \label{sec:arch} \input{arch}

\section{RemIX Deployed in Scotland} \label{sec:impl} \input{scotland}

%\section{Guifi in Catalonia} \label{sec:impl} \input{catalonia}

\section{The Environment}\label{sec:hubs}\input{hubs}

\section{Concluding Remarks} \label{sec:conc} \input{conc}

\section{Acknowledgements}
This project is supported by the University of Edinburgh, by the
Scottish Government and by local industries including Marine Harvest
Scotland and Benchmark Holdings.  During an incubation period,
\ac{WHIX} is jointly operated by HUBS and the University of Edinburgh.

\section{Additional Authors}
Additional authors: Michael Fourman (University of
Edinburgh), Richard Simmons (University of Stirling).

\bibliographystyle{abbrv}
\bibliography{paper_cix}

\end{document}
